---
title: "Presentation"
author: "Minsu Kim, Alona Muzikansky, Ariane Stark, Jadey Wu"
output: 
  beamer_presentation:
    theme: "Madrid"

---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(gtable)
library(gridExtra)
library(DescTools)
library(tidyverse)
library(kableExtra)
library(knitr)
library(tidyverse)
library(vcdExtra, quietly = TRUE)
library(ResourceSelection)
library(nnet)
```


```{r include=FALSE}
data <- read.csv("Data/initial_table.csv")
#mean(data$AGE, na.rm = T)
data.work <- dplyr::select(data, ID, AGE, SEX, IBS_POST, DLIT_AG, SIM_GIPERT, endocr_01, endocr_02, ZSN, LET_IS) 
data.work <- na.omit(data.work)
data.work <- filter(data.work, DLIT_AG != 10)
```


## Alona: Examining the relationship between Duration of Arterial Hypertension and CHF

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=5 ,paged.print=FALSE, fig.align = "center"}
freqplot <- ggplot(data.work, aes(x = as.factor(ZSN), y = as.factor(DLIT_AG))) +
  geom_count() +
  labs(title = "", x = "", y="") + 
  scale_x_discrete(labels = c("No", "Yes")) +
  scale_y_discrete(labels = c("None","1","2","3","4","5","6-10",">=10")) 
freqplot
```

## 
* The two classes of CHF have similar count distributions across the levels of duration of arterial hypertension. 

* We will further test the hypothesis that there is an association between the two variables

## Inference for contigency table.
```{r echo=FALSE, message=FALSE, warning=FALSE}

data.work.2 <- data.work %>% 
  mutate(DLIT_AG_N = case_when(DLIT_AG==6 ~ 8,  
                               DLIT_AG==7 ~ 10,
                               DLIT_AG==1 ~ 1,
                               DLIT_AG==2 ~ 2,
                               DLIT_AG==3 ~ 3,
                               DLIT_AG==4 ~ 4,
                               DLIT_AG==5 ~ 5,
                               DLIT_AG==0 ~ 0
                               ))

tab <- table(data.work.2$DLIT_AG,data.work.2$ZSN)
dimnames(tab) <- list("Duration of AH"=c("None","1","2","3","4",
                                         "5","6-10",">=10"),
                      "Chronic Heart Failure"=c("No","Yes"))

# contingency table
dlitag <- as.table(tab) 
kable(dlitag,
      caption = "Duration of Arterial Hypertension by Chronic Heart Failure")

```



## Examining the Standerdized residuals.
```{r ,echo=FALSE, message=FALSE, warning=FALSE}
# chisq test can be used but is less powerful than the two above.
chisq <- round(chisq.test(dlitag)$statistic,3)
#pval <- round(chisq.test(dlitag)$p.value,3)
#lrt <- GTest(dlitag)
std.res <- chisq.test(dlitag)$stdres

# residual analysis
mosaicplot(dlitag,
           main = "",
           xlab = "Duration of Arterial Hypertension",
           ylab = "Chronic Heart Failure",
           las = 1,
           border = "chocolate",
           shade = TRUE)

```

## 
For Ix2 tables, testing for a linear trend in either response category, we use the Cochran-Armitage trend test. 

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
# Cochran Armitage Test for Ix2 tables - section 5.3.5 in the book
coarm <- CochranArmitageTest(dlitag)
coarm
```

Issues to consider: Ordinal variable with unequal intervals so trend test on the original classification provides information about the direction but ignores the unequal spacing in the last two categories.


## Logistic Regression model
x - Duration of Arterial Hypertension.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
x=c(0,1,2,3,4,5,8,10)
yes=c(120,21,10,12,4,20,37,104)
no=c(401,72,47,42,15,48,120,307)
plotdata <- data.frame(x,yes, no)

model.logit <- glm(cbind(yes, no) ~ x, family = binomial("logit"))
infologit <- summary(model.logit)
model.linear <- glm(cbind(yes, no) ~ x, family = binomial("identity"))
infolinear <- summary(model.linear)
model.probit <- glm(cbind(yes, no) ~ x, family = binomial("probit")) 
infoprobit <- summary(model.probit)
model.cloglog <- glm(cbind(yes, no) ~ x, family = binomial("cloglog"))
infocloglog <- summary(model.cloglog)

y_logit <- predict(model.logit, list(xval = plotdata$x), type="response")
y_linear <- predict(model.linear, list(xval = plotdata$x), type="response")
y_probit <- predict(model.probit, list(xval = plotdata$x), type="response")
y_cloglog <- predict(model.cloglog, list(xval = plotdata$x), type="response")

kable(infologit$coef, caption="Parameter Estimates for Logit link")
kable(infolinear$coef, caption="Parameter Estiamtes for Identity link")
#kable(modelinfo.cloglog$coef, caption="Parameter Estiamtes for cloglog link")
```

## Goodness of fit tests for the fitted models

For the logit  model:
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
G.sq=deviance(model.logit)
df.fit <- model.logit$df.residual
p.val=1-pchisq(G.sq,df.fit)
```
* $G^2$ = `r I(G.sq)`
* df = `r I(df.fit)`
* p-value = `r I(p.val)`

For the linear model:
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
G.sq=deviance(model.linear)
df.fit <- model.linear$df.residual
p.val=1-pchisq(G.sq,df.fit)
```
* $G^2$ = `r I(G.sq)`
* df = `r I(df.fit)`
* p-value = `r I(p.val)`

## 
Predicted probabilities for the fitted models and the observed data.
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(plotdata$x, plotdata$yes/(plotdata$yes +plotdata$no),
     xlab="Duration of arterial hypertension", 
     ylab="Probability of CHF", 
     type = "p", pch=21, bg="black")
points(plotdata$x, y_logit, pch=21, bg="blue")
lines(plotdata$x, y_linear , col=2)
#points(plotdata$x, y_probit , col=5)
#points(plotdata$x, y_cloglog , col=5)
legend("bottomright", legend=c("Logistic - Blue dots", "Linear - Red line","Observed data - Black dots"))
```


## Sub-analsis
We tested the Linear model for the subset: Duration of arterial hypertension $\in[1 - 5]$
 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
data.work.sub <- data.work.2 %>% 
  filter(DLIT_AG_N %in% c(1,2,3,4,5))
model <- glm(ZSN~DLIT_AG_N, data = data.work.sub, family = binomial("identity"))
model.info <- summary(model)
G.sq=deviance(model)
df.fit <- model$df.residual
p.val.sub=round((1-pchisq(G.sq,df.fit)),3)
kable(model.info$coef, caption="Parameter Estiamtes for subset analysis")
```

## 
Predicted probabilities
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, paged.print=FALSE}
newdata <- data.frame(DLIT_AG_N=seq(min(data.work.sub$DLIT_AG_N), max(data.work.sub$DLIT_AG_N),len=23))
newdata$ZSN <- predict(model, newdata=newdata, type="response")
plot(ZSN~DLIT_AG_N, data=data.work.sub, col="black",
     main = "Plot A",
     ylab = "Predicted probability of CHF", 
     xlab = "Duration of arterial hypertension")
lines(ZSN~DLIT_AG_N, newdata, col="Blue", lwd=2)
```
The p-value for the goodness of fit went down sharply (`r I(p.val.sub)`) but still didn't reach significance level to reject the null of no-fit. 


## Conclusions

* There is no significant association between CHF and the duration of arterial hypertension.
* By itself, duration of arterial hypertension is not predictive of CHF.



