---
title: "Chronic Heart Failure and Risk Factors in Myochardial Infarction Dataset"
author: "Ariane, Alona, Minsu, and Jadey"
output: 
   beamer_presentation:
    theme: "Madrid"
    slide_level: 2
    toc: FALSE

---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(gtable)
library(gridExtra)
library(DescTools)
library(tidyverse)
library(kableExtra)
library(knitr)
library(tidyverse)
library(vcdExtra, quietly = TRUE)
library(ResourceSelection)
library(nnet)
```


```{r include=FALSE}
data <- read.csv("Data/initial_table.csv")
#mean(data$AGE, na.rm = T)
data.work <- dplyr::select(data, ID, AGE, SEX, IBS_POST, DLIT_AG, SIM_GIPERT, endocr_01, endocr_02, ZSN, LET_IS) 
data.work <- na.omit(data.work)
data.work <- filter(data.work, DLIT_AG != 10)
```

# Introduction

## Chronic heart failure (CHF)  

According to CDC,

\ \ 

\begin{columns}[onlytextwidth,T]
  \begin{column}{.45\linewidth}
  \begin{itemize}
    \item More than 6 million adults in the USA have heart failure.
  \end{itemize}
  \end{column}
  \begin{column}{.45\linewidth}
    \includegraphics[width=0.65\linewidth]{chf.jpg}
  \end{column}
\end{columns}

\ \ 

 \begin{columns}[onlytextwidth,T]
  \begin{column}{.45\linewidth}
  \begin{itemize}
    \item About half of Americans (47\%) have at least one of key risk factors.
  \end{itemize}
  \end{column}
  \begin{column}{.45\linewidth}
    \includegraphics[width=0.6\linewidth]{causes.jpg}
  \end{column}
\end{columns}

\begingroup
\fontsize{5}{7}\selectfont
(Figure(up): https://www.disability-benefits-help.org/resources/medical-evidence/chronic-heart-failure)
(Figure(down):https://www.verywellhealth.com/heart-failure-causes-and-risk-factors-1746181)
\endgroup


## Topics to be covered

**Question:** How are the predictors of our interest associated with Chronic heart failure?

- Test independence of demographics with regards to CHF

- Association of duration of arterial hypertension and CHF

- Build a multiple logistic regression model by adding more predictors and identify the best model

- Modeling the relationship between death outcome and selected variables

# Dataset Overview

## Descriptive statistics

- Demographic information
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align="center",fig.height=5.75}
# distribution plots for single variable

age.hist <- ggplot(data.work, aes(data.work$AGE)) + geom_histogram() + 
  labs(title = "age distribution", x = "age") + 
  geom_vline(xintercept = mean(data.work$AGE), color = "blue") + 
  geom_vline(xintercept = median(data.work$AGE), linetype = "dotted") +
  annotate("text", x = 40, y = 120, label = "mean = 61.397") +
  annotate("text", x = 40, y = 110, label = "median = 62") 
#age.hist

sex.plot <- ggplot(data.work, aes(as.factor(data.work$SEX))) + geom_bar() + 
  labs(title = "distribution of sex", x = "sex", y = "") + 
  scale_x_discrete(labels = c("female", "male")) +
  annotate("text", x = 1, y = 780, label = "female = 502") +
  annotate("text", x = 1, y = 720, label = "male = 878") 
#sex.plot

grid.arrange(age.hist, sex.plot, nrow = 1)

```

## Descriptive statistics
- Patient physiological attributes
- IBS_POST: coronary heart disease in recent weeks before admission to hospital
  - 0: there was no CHD
  - 1: extertional angina pectoris
  - 2: unstable angina pectoris

- DLIT_AG: duration of arterial hypertension
  - 0: there was no arterial hypertension
  - 1: one year
  - 2: two years
  - 3: three years
  - 4: four years
  - 5: five years
  - 6: 6-10 years
  - 7: more than 10 years
  
## Descriptive statistics  
- SIM_GIPERT: systematic hypertension; 0 - no, 1 - yes
- endocr_01: diabetes mellitus in the anamnesis; 0 - no, 1 - yes
- endocr_02: obesity in the anamnesis; 0 - no, 1 - yes
- ZSN: chronic heart failure; 0 - no, 1 - yes

## Descriptive statistics

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align = "center", fig.height=5.75}
ibs.plot <- ggplot(data.work, aes(as.factor(data.work$IBS_POST))) + 
  geom_bar() + 
  labs(title = "IBS_POST", x = "type of CHD") +
  annotate("text", x = 1, y = 560, label = "0: 353") +
  annotate("text", x = 1, y = 500, label = "1: 443") +
  annotate("text", x = 1, y = 440, label = "2: 584") 
#ibs.plot

duration.hist <- ggplot(data.work, aes(data.work$DLIT_AG)) + 
  geom_histogram() + 
  labs(title = "DLIT_AG", x = "years", y = "")
#duration.hist

hypertension.plot <- ggplot(data.work, aes(as.factor(data.work$SIM_GIPERT))) + 
  geom_bar() + 
  labs(title = "SIM_GIPERT", x = "hypertension", y = "") + 
  scale_x_discrete(labels = c("no", "yes")) + 
  annotate("text", x = 2, y = 1100, label = "no: 1336") +
  annotate("text", x = 2, y = 900, label = "yes: 44") 
#hypertension.plot

diabetes.plot <- ggplot(data.work, aes(as.factor(data.work$endocr_01))) + 
  geom_bar() + 
  labs(title = "endocr_01", x = "diabetes") + 
  scale_x_discrete(labels = c("no", "yes")) +
  annotate("text", x = 2, y = 1100, label = "no: 1193") +
  annotate("text", x = 2, y = 900, label = "yes: 187") 
#diabetes.plot

obesity.plot <- ggplot(data.work, aes(as.factor(data.work$endocr_02))) + 
  geom_bar() + 
  labs(title = "endocr_02", x = "obesity", y = "") + 
  scale_x_discrete(labels = c("no", "yes")) +
  annotate("text", x = 2, y = 1100, label = "no: 1348") +
  annotate("text", x = 2, y = 900, label = "yes: 32") 
#obesity.plot

chd.plot <- ggplot(data.work, aes(as.factor(data.work$ZSN))) + 
  geom_bar() + 
  labs(title = "ZSN", x = "CHF", y = "") + 
  scale_x_discrete(labels = c("no", "yes")) +
  annotate("text", x = 2, y = 950, label = "no: 1052") +
  annotate("text", x = 2, y = 800, label = "yes: 328") 
#chd.plot

grid.arrange(ibs.plot, duration.hist, hypertension.plot, diabetes.plot, obesity.plot, chd.plot, nrow = 2)
```

# Tests for Independence of Demographics

## Analysis of Sex and Chronic Heart Failure: Overview

Question: Is there an association between sex and chronic heart failure?

```{r echo=FALSE,comment=NA }
data_sex_chf <- table(data.work$SEX,data.work$ZSN)
dimnames(data_sex_chf) <- list(Sex=c("Female","Male"),         
                       "Chronic Heart Failure"=c("No","Yes"))
data_sex_chf
```

## Analysis of Sex and Chronic Heart Failure: Tests

Pearson $\chi^2$ Test of Independence:

```{r echo=FALSE,  comment=NA}
chi_sq_data_sex_chf <-chisq.test(data_sex_chf)
chi_sq_data_sex_chf$statistic
cat(paste("p-value =",round(chi_sq_data_sex_chf$p.value,5)))
```

Likelihood Ratio Test of Independence:

```{r echo=FALSE,comment=NA}
LR_data_sex_chf <- GTest(data_sex_chf)
LR_data_sex_chf$statistic
cat(paste("p-value =",round(LR_data_sex_chf$p.value,5)))
```
## Analysis of Age(Continuous) and Chronic Heart Failure: Overview

Question: Is there an association between age and chronic heart failure?
```{r echo=FALSE, fig.height=5.75, comment=NA}
# age and chronic heart failure
data_age_chf <- table(data.work$AGE,data.work$ZSN)
dimnames(data_age_chf) <- list(Age = names(data_age_chf[,1]),
                               "Chronic Heart Failure"=c("No","Yes"))
#data_age_chf

boxplot_age_chf <- data.work %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x=AGE, y=as.factor(ZSN), 
                             group = as.factor(ZSN))) +
  ylab("Chronic Heart Failure") +
  scale_y_discrete(labels=c("No","Yes"))
boxplot_age_chf
```

## Analysis of Age(Continuous) and Chronic Heart Failure: Summary Statistics

```{r echo=FALSE,comment=NA}
#CHF NO
chf_no <- summary(data.work %>% 
          filter(ZSN==0) %>% 
          select(AGE))

#CHF YES
chf_yes <- summary(data.work %>% 
          filter(ZSN==1) %>% 
          select(AGE))

chf_age_cont_descriptive_statistcs <- data.work %>% 
  transform("ZSN" = ifelse(ZSN==1,"Yes","No")) %>% 
  group_by(ZSN) %>% 
  summarise_at(vars("AGE"), list( "Min." = min,
            "1st Qu." = ~quantile(.,prob=0.25),
            "Median" = ~quantile(.,prob=0.5),
            "Mean" = mean,
            "3rd Qu." = ~quantile(.,prob=0.75),
            "Max" = max)) 

chf_age_cont_descriptive_statistcs_t <-
  as.table(t(as.matrix.data.frame(chf_age_cont_descriptive_statistcs)))[-1,]


dimnames(chf_age_cont_descriptive_statistcs_t) <- list(" "=rownames(chf_age_cont_descriptive_statistcs_t),         
                       "Chronic Heart Failure"=c("No","Yes"))

print(chf_age_cont_descriptive_statistcs_t)
```

## Analysis of Age(Continuous) and Chronic Heart Failure: Test

Analysis was done using a two sided Wilcoxon Rank Sum Test to test if there is a difference in Chronic Heart Failure outcome across age.

```{r echo=FALSE,comment=NA}
wilcox_test <- wilcox.test(data.work$AGE[which(data.work$ZSN == 0)],
            data.work$AGE[which(data.work$ZSN == 1)])
wilcox_test$statistic
cat(paste("p-value =",round(wilcox_test$p.value,8)))
```
## Analysis of Age(Categorical) and Chronic Heart Failure: Overview

Question: Is there an association between age(decade) and chronic heart failure?
```{r echo=FALSE,comment=NA}
age_decade <- data.work %>% 
  mutate(decade = floor(AGE/10)*10) %>% 
  select(decade)
data_age_decade_chf <- table(age_decade$decade,data.work$ZSN)
dimnames(data_age_decade_chf) <- 
  list(Age = paste0(names(data_age_decade_chf[,1]),"s"),
                               "Chronic Heart Failure"=c("No","Yes"))
data_age_decade_chf
```

## Analysis of Age(Categorical) and Chronic Heart Failure: Test

Pearson $\chi^2$ Test of Independence:
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
chi_sq_data_age_decade_chf <-chisq.test(data_age_decade_chf)
chi_sq_data_age_decade_chf$statistic
cat(paste("p-value =",round(chi_sq_data_age_decade_chf$p.value,5)))
```

Likelihood Ratio Test of Independence:
```{r echo=FALSE,comment=NA}
LR_data_age_decade_chf <- GTest(data_age_decade_chf)
LR_data_age_decade_chf$statistic
cat(paste("p-value =",round(LR_data_age_decade_chf$p.value,8)))
```

# Association of duration of arterial hypertension and CHF

## Examining the relationship between Duration of Arterial Hypertension and CHF

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=5 ,paged.print=FALSE, fig.align = "center"}
freqplot <- ggplot(data.work, aes(x = as.factor(ZSN), y = as.factor(DLIT_AG))) +
  geom_count() +
  labs(title = "", x = "", y="") + 
  scale_x_discrete(labels = c("No", "Yes")) +
  scale_y_discrete(labels = c("None","1","2","3","4","5","6-10",">=10")) 
freqplot
```

## 
* The two classes of CHF have similar count distributions across the levels of duration of arterial hypertension. 

* We will further test the hypothesis that there is an association between the two variables

## Inference for contigency table.
```{r echo=FALSE, message=FALSE, warning=FALSE}

data.work.2 <- data.work %>% 
  mutate(DLIT_AG_N = case_when(DLIT_AG==6 ~ 8,  
                               DLIT_AG==7 ~ 10,
                               DLIT_AG==1 ~ 1,
                               DLIT_AG==2 ~ 2,
                               DLIT_AG==3 ~ 3,
                               DLIT_AG==4 ~ 4,
                               DLIT_AG==5 ~ 5,
                               DLIT_AG==0 ~ 0
                               ))

tab <- table(data.work.2$DLIT_AG,data.work.2$ZSN)
dimnames(tab) <- list("Duration of AH"=c("None","1","2","3","4",
                                         "5","6-10",">=10"),
                      "Chronic Heart Failure"=c("No","Yes"))

# contingency table
dlitag <- as.table(tab) 
kable(dlitag,
      caption = "Duration of Arterial Hypertension by Chronic Heart Failure")

```



## Examining the Standerdized residuals.
```{r ,echo=FALSE, message=FALSE, warning=FALSE}
# chisq test can be used but is less powerful than the two above.
chisq <- round(chisq.test(dlitag)$statistic,3)
#pval <- round(chisq.test(dlitag)$p.value,3)
#lrt <- GTest(dlitag)
std.res <- chisq.test(dlitag)$stdres

# residual analysis
mosaicplot(dlitag,
           main = "",
           xlab = "Duration of Arterial Hypertension",
           ylab = "Chronic Heart Failure",
           las = 1,
           border = "chocolate",
           shade = TRUE)

```

## 
For Ix2 tables, testing for a linear trend in either response category, we use the Cochran-Armitage trend test. 

```{r ,echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
# Cochran Armitage Test for Ix2 tables - section 5.3.5 in the book
coarm <- CochranArmitageTest(dlitag)
coarm
```

Issues to consider: Ordinal variable with unequal intervals so trend test on the original classification provides information about the direction but ignores the unequal spacing in the last two categories.


## Logistic Regression model
x - Duration of Arterial Hypertension.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
x=c(0,1,2,3,4,5,8,10)
yes=c(120,21,10,12,4,20,37,104)
no=c(401,72,47,42,15,48,120,307)
plotdata <- data.frame(x,yes, no)

model.logit <- glm(cbind(yes, no) ~ x, family = binomial("logit"))
infologit <- summary(model.logit)
model.linear <- glm(cbind(yes, no) ~ x, family = binomial("identity"))
infolinear <- summary(model.linear)
model.probit <- glm(cbind(yes, no) ~ x, family = binomial("probit")) 
infoprobit <- summary(model.probit)
model.cloglog <- glm(cbind(yes, no) ~ x, family = binomial("cloglog"))
infocloglog <- summary(model.cloglog)

y_logit <- predict(model.logit, list(xval = plotdata$x), type="response")
y_linear <- predict(model.linear, list(xval = plotdata$x), type="response")
y_probit <- predict(model.probit, list(xval = plotdata$x), type="response")
y_cloglog <- predict(model.cloglog, list(xval = plotdata$x), type="response")

kable(infologit$coef, caption="Parameter Estimates for Logit link")
kable(infolinear$coef, caption="Parameter Estiamtes for Identity link")
#kable(modelinfo.cloglog$coef, caption="Parameter Estiamtes for cloglog link")
```



## 
Predicted probabilities for the fitted models and the observed data.
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(plotdata$x, plotdata$yes/(plotdata$yes +plotdata$no),
     xlab="Duration of arterial hypertension", 
     ylab="Probability of CHF", 
     type = "p", pch=21, bg="black")
points(plotdata$x, y_logit, pch=21, bg="lightgreen")
lines(plotdata$x, y_linear , col=2)
#points(plotdata$x, y_probit , col=5)
#points(plotdata$x, y_cloglog , col=5)
legend("bottomright", legend=c("Logistic - Blue dots", "Linear - Red line","Observed data - Black dots"))
```


## Sub-analysis
We tested the Linear probability model for the subset: Duration of arterial hypertension between 1 and 5.
 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
data.work.sub <- data.work.2 %>% 
  filter(DLIT_AG_N %in% c(1,2,3,4,5))
model <- glm(ZSN~DLIT_AG_N, data = data.work.sub, family = binomial("identity"))
model.info <- summary(model)
G.sq=deviance(model)
df.fit <- model$df.residual
p.val.sub=round((1-pchisq(G.sq,df.fit)),3)
kable(model.info$coef, caption="Parameter Estiamtes for subset analysis")
```




# Multiple Logistic Regression and Model Selection

## Multiple Logistic Regression

- Coefficient estimates of the multiple logistic regressions of all predictors
```{r echo=FALSE }
data.work <- data.work %>% 
  mutate(DLIT_AG_N = case_when(DLIT_AG==6 ~ 8,  
                               DLIT_AG==7 ~ 10,
                               DLIT_AG==0 ~ 0,
                               DLIT_AG==1 ~ 1,
                               DLIT_AG==2 ~ 2,
                               DLIT_AG==3 ~ 3,
                               DLIT_AG==4 ~ 4,
                               DLIT_AG==5 ~ 5
                               ))
data.work$SIM.f <- factor(data.work$SIM_GIPERT, levels=c(0,1), labels = c("no","yes"))
data.work$endocr_01.f <- factor(data.work$endocr_01, levels=c(0,1), labels = c("no","yes"))
data.work$endocr_02.f <- factor(data.work$endocr_02, levels=c(0,1), labels = c("no","yes"))
chf.dat <- select(data.work, AGE, SEX, IBS_POST, DLIT_AG_N, SIM.f, endocr_01.f, endocr_02.f, ZSN) 
fit<- glm(ZSN ~ . , data=chf.dat, family=binomial)
kbl(summary(fit)$coefficients, digits = 5) 
```

- Only AGE and endocr_01 are statistically significant.
- The P-value for the overall test is much less than 0.0001, thus there is strong evidence that at least one predictor has an effect.

## Multiple Logistic Regression - Goodness of Fit

Fit a multiple logistic regression model by adding AGE and endocr_01 to the logistic regression model with only DLIT_AG:
$$\text{logit}[P(ZSN=1)]=\alpha+\beta_1\,DLIT\_AG+\beta_2\,AGE+\beta_3\,endocr\_01.f.$$

- Goodness of Fit

```{r echo=FALSE}
fit.ini <- glm(ZSN~ DLIT_AG_N, data=chf.dat, family=binomial)
fit.add <- glm(ZSN~ DLIT_AG_N + AGE + endocr_01.f, data=chf.dat, family=binomial)

G.sq=deviance(fit.add)
df <- fit.add$df.residual
p.value=1-pchisq(G.sq,df)

tab1 <- c(G.sq, df, p.value)
names(tab1) <- c("G.square", "df", "P-value")
kable(t(tab1), escape=FALSE)
```

The model has $G^2=1459$ with degree of freedom $df = 1376$ (P-value= 0.059 > 0.05), which indicates a decent fit. 

## Multiple Logistic Regression - ANOVA test

Comparing this additive model with the initial model with DLIT_AG only, 

- ANOVA Result

```{r echo=FALSE}
kable(anova(fit.ini, fit.add, test="LRT"), digits = 2)
```
the likelihood ratios test statistic is 53.73 with degree of freedom 2, producing very tiny p-value (P<0.001). Thus, the model with AGE and endocr_01 in addition to DLIT_AG improves the goodness-of-fit.


## Multiple Logistic Regression - Model selection
We perform stepwise model selection to see if there is effect of interaction between predictors.  

- Backward selection

```{r echo = FALSE}
#Backward
fit.3 <- glm(ZSN~ DLIT_AG_N* AGE * endocr_01.f, data=chf.dat, family=binomial)
mod.back <- step(fit.3, scope=list(lower = ~ 1, upper = formula(fit.3)), 
                 scale = 1, trace = F, direction = "backward")
res.back <- mod.back$anova
kable(res.back, digits = 3) %>% kable_styling(font_size = 8)
```

## Multiple Logistic Regression - Model selection

- Forward selection

```{r echo = FALSE}
#Forward
fit.0 <- glm(ZSN ~ 1 , data=chf.dat, family=binomial)
mod.for <- step(fit.0, scope=list(lower = ~ 1, upper = formula(fit.3)), 
                scale = 1, trace = F, direction = "forward")
res.for <- mod.for$anova
kable(res.for, digits = 3)%>% kable_styling(font_size = 8)
```

Based on the AIC, both backward elimination and forward selection choose the model of 
\begin{align*}
\text{logit}[P(ZSN=1)]=&\alpha+\beta_1\,DLIT\_AG+\beta_2\,AGE+\beta_3\,endocr\_01.f \\
 &+\beta_4\,DLIT\_AG * endocr_01.f.
\end{align*}

## Predictive Power - ROC curves

- ROC curves of the selected model with interaction and the additive model 

```{r echo=FALSE, fig.cap = "ROC curves\\label{fig:rocplot}", fig.align='center', out.width = '80%'}
fit.best <- glm(ZSN ~ AGE + DLIT_AG_N * endocr_01.f , data=chf.dat, family=binomial)
library(ROCR)
pred1 <- prediction(fitted(fit.add), chf.dat$ZSN)
stats1 <- performance(pred1, 'tpr', 'fpr')

pred2 <- prediction(fitted(fit.best), chf.dat$ZSN)
stats2 <- performance(pred2, 'tpr', 'fpr')

mod1.lab <- expression('AGE'+'DLIT_AG'+'endocr_01')
mod2.lab <- expression('AGE'+'DLIT_AG x endocr_01')

plot(stats1@x.values[[1]], stats1@y.values[[1]], type='s', 
     ylab=stats1@y.name, xlab=stats1@x.name, col='red', lwd=2, lty=2)
lines(stats2@x.values[[1]], stats2@y.values[[1]], type='s', col='blue', lwd=2, )
abline(0,1, col='gray')
legend('bottomright', c(mod1.lab, mod2.lab), col=c('red','blue'), 
       lwd=c(2,1), lty=c(2,1), cex=.9, bty='n')
```

Their ROC curves are very close, thus we expect their performance for prediction will be almost same. 

# Modeling the relationship between death outcome and selected variables 

## Secondary analysis  
- The dataset includes one variable indicating the causes of lethal outcome for the patients
  - LET_IS: causes of lethal outcome
    - 0: survive
    - 1: cardiogenic edema
    - 2: pulmonary edema
    - 3: myocardial rupture
    - 4: progress of congestive heart failure
    - 5: thromboembolism
    - 6: asystole
    - 7: ventricular fibrillation
- Build a logistic regression model to predict death of the patients by turning LET_IS to a binary variable "death"
- Build model with multimonial response to investigate the cause of death

## Secondary analysis
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=5.75}
data.work2 <- data.work
data.work2$death <- ifelse(data.work$LET_IS == 0, 0, 1)
#table(data.work2$death) # survive: 1212, dead: 191

death.plot <- ggplot(data.work, aes(as.factor(data.work2$death))) + 
  geom_bar() + 
  labs(title = "distribution of death", x = "death") + 
  scale_x_discrete(labels = c("no", "yes")) +
  annotate("text", x = 2, y = 1000, label = "no: 1212") +
  annotate("text", x = 2, y = 800, label = "yes: 191")

data.work3 <- filter(data.work2, LET_IS != 0)
#dim(data.work3) # n = 191
#table(data.work3$LET_IS)
cause.plot <- ggplot(data.work3, aes(as.factor(data.work3$LET_IS))) + 
  geom_bar() + 
  labs(title = "causes of death") + 
  xlab("") + 
  scale_x_discrete(labels = c("cardiogenic shock", "pulmonary endema", "myocardial rupture", "progress of conestive heart failure", "thromboembolism", "asystole", "ventricular fibrillation"), guide = guide_axis(n.dodge = 2))

grid.arrange(death.plot, cause.plot, nrow = 2)
```


## Secondary analysis
- Full model contains continuous variables age, duration of arterial hypertension, categorical variables SEX, chronic heart disease duration before admission to hospital, systematic hypertension, diabetes, obesity, and the interaction terms between AGE and all the other variables.
- Used stepwise step() to select the best model.
- The best model selected:

$log[P(death = 1)] = \beta_0 + \beta_1 \times age + \beta_3 \times I(IBS = 1) + \beta_4 \times I(IBS = 2) + \beta_5 \times I(SIM = 1) + beta_6 \times I(endocr01 = 1) + beta_7 \times I(endocr02 = 1)$

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# use stepwise selection to select variable
death.fit0 <- glm(death ~ 1, data = data.work2, family = binomial)
death.fit1 <- glm(death ~ AGE + as.factor(SEX) + as.factor(IBS_POST) + DLIT_AG + as.factor(SIM_GIPERT) + as.factor(endocr_01) + as.factor(endocr_02) + AGE * factor(IBS_POST) + AGE * DLIT_AG + AGE * factor(SIM_GIPERT) + AGE * factor(endocr_01) + AGE * factor(endocr_02), data = data.work2, family = binomial)
death.mod <- step(death.fit1, death.fit0, direction = "both", trace = F) # selected variable: AGE, IBS_POST, SIM_GIPERT, endocr_01, endocr_02
summary(death.mod)

# fit the best model 
death.fit.logit <- glm(death ~ AGE + as.factor(IBS_POST) + as.factor(SIM_GIPERT) + as.factor(endocr_01) + as.factor(endocr_02), data = data.work2, family = binomial)
summary(death.fit.logit)
deviance(death.fit.logit) # 1018.906

exp(death.fit.logit$coefficients)
```

```{r}
kable(summary(death.mod)$coefficients, digits = 3) %>% kable_styling(font_size = 8)
```


## Secondary analysis

- Goodness of fit check with Hosmer-Lemeshow test by grouping the observations into 20 groups. The test statistic is 0.4291, indicating an adequate fit of the model to the dataset.
- Plotted the predicted value against the observed value of the 20 groups. Overall the dots follow the diagonal. 

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# Hosmer-Lemeshow test to check goodness of fit
library("ResourceSelection")
death.pred <- predict(death.fit.logit, data.work2, type = "response")
hoslem.test(data.work2$death, death.pred, g = 20) # p = 0.4291, fail to reject H0
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
## Get indices of vector fit, from smallest to greatest
fit <- death.fit.logit$fitted.values
index <- sort.list(fit)
## check 10 smallest indices
index[1:10]
## create a matrix of death and fit, using this index
hosmer <- matrix(c(data.work2$death[index], fit[index]), byrow = F, nrow = nrow(data.work2))
head(hosmer)
## group into 20 groups with 69 observations per group
observed <- rep(NA, 20)
for (i in 1:20){ observed[i] <- sum(hosmer[(69*(i-1) +1) : (69 *i), 1])/ 69 }
observed
# repeat the previous step for the predicted probability
predicted <- rep(NA, 20)
for (i in 1:20){ predicted[i] <- sum(hosmer[(69*(i-1) +1) : (69 *i), 2])/ 69 }
predicted
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width = "200px", out.height="300px", fig.align = "center"}
# plot observed versus predicted
ggplot() + aes(x = predicted, y = observed) + geom_point() + geom_line() + geom_abline( a = 0, b = 1, color = "red")
```


## Secondary analysis
- Fit baseline category logit model on cause of death. Used predictors selected in the previous analysis. 
$log \frac{\pi_j(x)}{\pi_J(x)} = \beta_{0j} + \beta_{1j} \times age + \beta_{2j} \times I(IBS = 1) + \beta_{3j} \times I(IBS = 2) + \beta_{4j} \times I(SIM = 1) + \beta_{5j} \times I(endocr01 = 1) + \beta_{6j} \times I(endocr02 = 1), j = 1, ..., 6$

where J = cardiogenic shock, j = 1 pulmonary edema, 2 myocardial rupture, 3 progress of congestive heart failure, 4 thromboembolism, 5 asystole, 6 ventricular fibrillation

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
multi.mod <- multinom(LET_IS ~ AGE + as.factor(IBS_POST) + as.factor(SIM_GIPERT) + as.factor(endocr_01) + as.factor(endocr_02), data = data.work3)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, out.width = "200px", out.height="300px", fig.align="cente"}
output <- summary(multi.mod)
table <- as.data.frame(output$coefficients)
colnames(table) <- c("intercept", "AGE", "IBS_POST = 1", "IBS_POST = 2", "SIM_GIPERT = 1", "endocr_01 = 1", "endocr_02 = 1")
```

## Secondary analysis
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
kable(table, digits = 3)  %>% kable_styling(font_size = 6)
```


# Conclusion

## 

- Age and Sex are associated with CHF
- Duration of Arterial Hypertension is predictive when included in a multivariate model
- The final multivariable model for CHF is not rejected
- Age, coranary heart disease in recent weeks, symptomatic hypertension, obesity and diabetes are associated with patient death.
