---
title: "Workfile"
author: "Group"
output: pdf_document
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(ggpubr)
```


```{r message=FALSE, warning=FALSE}
data <- read.csv("Data/initial_table.csv")
#mean(data$AGE, na.rm = T)
data.work <- select(data, ID, AGE, SEX, IBS_POST, DLIT_AG, SIM_GIPERT, endocr_01, endocr_02, ZSN, LET_IS) 
data.work <- na.omit(data.work)
dim(data.work) # obs = 1381 

# exploratory analysis
mean(data.work$AGE) #61.3997
table(data.work$SEX) # female(0): 502, male(1): 879
table(data.work$IBS_POST) # no CHD(0): 354, exertional angina pectoris(1):443, unstable angina pectoris(2):584
mean(data.work$DLIT_AG) # 3.3407
table(data.work$SIM_GIPERT) # no(0): 1337, yes(1): 44
table(data.work$endocr_01) # no(0): 1194, yes(1):187
table(data.work$endocr_02) # no(0): 1349, yes(1):32

# distribution plots for single variable
age.hist <- ggplot(data.work, aes(data.work$AGE)) + geom_histogram() + labs(title = "age distribution", x = "age")

sex.plot <- ggplot(data.work, aes(as.factor(data.work$SEX))) + geom_bar() + labs(title = "distribution of sex", x = "sex") + scale_x_discrete(labels = c("female", "male"))

ibs.plot <- ggplot(data.work, aes(as.factor(data.work$IBS_POST))) + geom_bar() + labs(title = "distribution of CHD in recent weeks", x = "type of CHD") + scale_x_discrete(labels = c("no CHD", "extertional angina pectoris", "unstable angina pectoris"))

duration.hist <- ggplot(data.work, aes(data.work$DLIT_AG)) + geom_histogram() + labs(title = "duration of arterial hypertension", x = "years")

hypertension.plot <- ggplot(data.work, aes(as.factor(data.work$SIM_GIPERT))) + geom_bar() + labs(title = "distribution of hypertension", x = "hypertension") + scale_x_discrete(labels = c("no", "yes"))

diabetes.plot <- ggplot(data.work, aes(as.factor(data.work$endocr_01))) + geom_bar() + labs(title = "distribution of diabetes", x = "diabetes") + scale_x_discrete(labels = c("no", "yes"))

obesity.plot <- ggplot(data.work, aes(as.factor(data.work$endocr_02))) + geom_bar() + labs(title = "distribution of obesity", x = "obesity") + scale_x_discrete(labels = c("no", "yes"))

ggarrange(age.hist, duration.hist, ncol = 2, nrow = 1)

ggarrange(sex.plot, ibs.plot, hypertension.plot, diabetes.plot, obesity.plot, ncol = 3, nrow = 2)

# plots of the relationship between CHD and single variable

```



```{r}
names(data)
```


## Ariane
```{r}
library("DescTools")
#sex and chronic heart failure

data_sex_chf <- table(data.work$SEX,data.work$ZSN)
dimnames(data_sex_chf) <- list(Sex=c("Female","Male"),         
                       "Chronic Heart Failure"=c("No","Yes"))
data_sex_chf
chi_sq_data_sex_chf <-chisq.test(data_sex_chf)
chi_sq_data_sex_chf

LR_data_sex_chf <- GTest(data_sex_chf)
LR_data_sex_chf
```


```{r}
# age and chronic heart failure
data_age_chf <- table(data.work$AGE,data.work$ZSN)
dimnames(data_age_chf) <- list(Age = names(data_age_chf[,1]),
                               "Chronic Heart Failure"=c("No","Yes"))
data_age_chf
chi_sq_data_age_chf <- chisq.test(data_age_chf)
chi_sq_data_age_chf

LR_data_age_chf <- GTest(data_age_chf)
LR_data_age_chf

```


 





## Alona
```{r}
#second chance-you did it
```


## Jadey
```{r}
data.work2 <- data.work
data.work2$death <- ifelse(data.work$LET_IS == 0, 0, 1)
table(data.work2$death) # survive: 1212, dead: 191

# fit logistic model

death.fit <- glm(death ~ . - ID - LET_IS - ZSN, data = data.work2, family = binomial)
summary(death.fit)

#death.fit2 <- glm(death ~ AGE + IBS_POST + endocr_02, data = data.work2, family = binomial)
#summary(death.fit2)

#anova(death.fit, death.fit2)
```


## Minsu
```{r}
data.work$SIM.f <- factor(data.work$SIM_GIPERT, levels=c(0,1), labels = c("no","yes"))
data.work$endocr_01.f <- factor(data.work$endocr_01, levels=c(0,1), labels = c("no","yes"))
data.work$endocr_02.f <- factor(data.work$endocr_02, levels=c(0,1), labels = c("no","yes"))
chf.dat <- select(data.work, AGE, SEX, IBS_POST, DLIT_AG, SIM.f, endocr_01.f, endocr_02.f, ZSN) 
fit<- glm(ZSN ~ . , data=chf.dat, family=binomial)
summary(fit)
```

We start by fitting a model containing all the 7 main effects,
$$\text{logit}[P(Y=1)]=\alpha+\beta_1AGE+\beta_2SEX+\beta_3IBS\_POST+\beta_4DLIT\_AG+\beta_5SIM+\beta_6endocr\_01+\beta_7ndocr\_02,$$
treating SIM_GIPERT, endocr_01, and endocr_01 as factors with indicator variables for "yes" for each attribute.  

```{r}
fit.0<- glm(ZSN ~ 1. , data=chf.dat, family=binomial)
summary(fit.0)
lr <- deviance(fit.0) - deviance(fit)
df <- summary(fit.0)$df[2]-summary(fit)$df[2]
p.val <- 1 - pchisq(lr, df=df)
p.val
```

The P-value for the overall test is much less than 0.0001, thus there is strong evidence that at least one predictor has an effect. Although the overall test is highly significant, Table 00 results shows that only AGE and endocr_01 seems significant in the Wald test. 

We now fit a multiple logistic regression model by adding AGE and endocr_01 to the logistic regression model in subtopic 2, which used DLIT_AG (Duration of arterial hypertension) alone as a predictor.   

```{r}
fit.ini <- glm(ZSN~ DLIT_AG, data=chf.dat, family=binomial)
fit.add <- glm(ZSN~ DLIT_AG + AGE + endocr_01.f, data=chf.dat, family=binomial)

```
```{r}
G.sq=deviance(fit.add)
df.fit <- fit.add$df.residual
p.val=1-pchisq(G.sq,df.fit)
```

The model has $G^2=1459.66$ with degree of freedom $df = 1377$ (P-value= 0.0597 > 0.05), which indicates a decent fit. 
Comparing this additive model with the initial model with DLIT_AG only, 

```{r}
anova(fit.ini, fit.add)
lr <- fit.ini$deviance - fit.add$deviance
lr
df <- anova(fit.ini, fit.add, test="LRT")$Df[2]
df
p.val <- 1 - pchisq(lr, df=df)
p.val
```


the likelihood ratios test statistic is 53.53 with degree of freedom 2, producing very tiny p-value (P<0.001). The model with AGE and endocr_01 in addition to DLIT_AG improves the goodness-of-fit. Next, we perform stepwise model selection through the forward and backward elimination methods to see if there is effect of interaction between predictors.  
```{r}
#Backward
fit.3 <- glm(ZSN~ DLIT_AG* AGE * endocr_01.f, data=chf.dat, family=binomial)
mod.back <- step(fit.3, scope=list(lower = ~ 1, upper = formula(fit.3)), scale = 1, trace = T, direction = "backward")
res.back <- mod.back$anova
res.back 
```

```{r}
#Forward
fit.0 <- glm(ZSN ~ 1 , data=chf.dat, family=binomial)
mod.for <- step(fit.0, scope=list(lower = ~ 1, upper = formula(fit.3)), scale = 1, trace = T, direction = "forward")
res.for <- mod.for$anova
res.for 
```
Based on the AIC, both backward elimination and forward selection choose the model AGE + DLIT_AG * endocr_01.f. 

```{r}
fit.best <- glm(ZSN ~ AGE + DLIT_AG * endocr_01.f , data=chf.dat, family=binomial)
summary(fit.best)
summary(fit.best)$coef
```

We can find that not only the added predictors, AGE and endocr_01, but also the interaction between DLIT_AG and endocr_01.f
are statistically significant, and  
```{r}
G.sq=deviance(fit.best)
df.fit.best <- fit.best$df.residual
p.val=1-pchisq(G.sq,df.fit.best)
```
this model with interaction term seems to fit well, as $G^2=1452.89$ with $df=1376$ (P-value = 0.073).

```{r}
lr <- fit.add$deviance - fit.best$deviance
lr
df <- anova(fit.ini, fit.best, test="LRT")$Df[2]
df
p.val <- 1 - pchisq(lr, df=df)
p.val
```

However, the likelihood ratios test statistic is 6.76 with df=3 (P-value=0.0799). The model with the interaction between AGE and DLIT_AG and endocr_01.f doesn't improve the goodness-of-fit.  
As seen in Figure , their ROC curves are very close, thus we expect their performance for prediction will be almost same. 

```{r}
library(ROCR)
pred1 <- prediction(fitted(fit.add), chf.dat$ZSN)
stats1 <- performance(pred1, 'tpr', 'fpr')

pred2 <- prediction(fitted(fit.best), chf.dat$ZSN)
stats2 <- performance(pred2, 'tpr', 'fpr')

mod1.lab <- expression('AGE'+'DLIT_AG'+'endocr_01')
mod2.lab <- expression('AGE'+'DLIT_AG x endocr_01')

plot(stats1@x.values[[1]], stats1@y.values[[1]], type='s', ylab=stats1@y.name, xlab=stats1@x.name, col='red', lwd=2)
lines(stats2@x.values[[1]], stats2@y.values[[1]], type='s', col='blue', lty=2)
abline(0,1, col='gray')
legend('bottomright', c(mod1.lab, mod2.lab), col=c('red','blue'), lwd=c(2,1), lty=1:2, cex=.9, bty='n')
```

